---
title: "miniproject"
author: "Cai Ytsma, Juliane Bombosh, Sofia Salcedo"
date: "3/19/2018"
output: html_document
  code_folding: hide
---
Notes for what to do:
    - What is the party of the people running for re-election? 
    - What is the proportion of people staying or new?
    - occupation of donors? Ranked occupations of democrats and republicans
        - group by party
        - Merge candidate and contribution, by contributor because then we can equal their id to their name and party
        - Then sort by party
        - Group by occupation 
        - Sum up the rows of each occupation
        - Arrange by the greatest number of rows
```{r}
library(tidyverse)
```

```{r}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


```{r}
contribution_code <- contributions %>%
  select(cand_id, state)

candidate_code <- candidates %>%
  mutate(party = cand_party_affiliation) %>%
  filter(party == "DEM" | party == "REP") %>%
  select(cand_id, party)
```

```{r}
contributions_party <- contribution_code %>%
  left_join(candidate_code, by = "cand_id") %>%
  group_by(party, state) %>%
  summarize(sum = n())
```

