---
title: "The States That Support Their Own"
subtitle: "Mini Project 2"
author: "Cai Ytsma, Juliane Bombosh, Sofia Salcedo"
date: "3/23/2018"
output: 
  html_document:
    code_folding: hide
---
# Relevance: 
**What we wanted to Achieve**
We analyzed FEC data to examine politically active local communities. We created two data graphics, one representing donations made to Democratic candidates, the other representing Republican candidates. We wanted to compare party affiliation and enthusiasm via donation frequency and examine the regions with higher political affiliation and why that might be.

#What We Did:
**Visual Cues & Data Source:**
Our data is from the candidates and contributions rda files. We utilized the image of a map and color saturation per state. Each state's saturation is based on the number of donations made to a candidate of that state. The higher the saturation the more donations were made to candidates of the repective party per state.

Where would donations to candidates be the highest? Would it be in areas where there is high variability in the local political scene and donations can have more influence? Or in areas that are the opposite, with a stable majority in place?
Donations can capture a glimpse of how politically active the community is in that moment. 

**The highest donations in-state correspond with swing states, reflecting a generally understood idea in American politics- money is power and money has the power to win elections.**




**Link to the repository** !!! NEEDS TO BE IN FOOTNOTES!!!

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(maps)
library(mapdata)
library(ggplot2)
library(devtools)
library(RColorBrewer)
library(extrafont)
```

```{r, label = "dataset import"}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, label = "state list"}
state_inits <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT",
               "DE", "FL", "GA", "HI", "ID",
               "IL", "IN", "IA", "KS", "KY", "LA", "ME",
               "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE",
               "NV", "NH", "NJ", "NM", "NY", "NC", "ND",
               "OH", "OK", "OR", "PA", "RI", "SC", 
               "SD", "TN", "TX", "UT", "VT", "VA", "WA", 
               "WV", "WI", "WY")

state_names <- c("alabama", "alaska", "arizona", "arkansas", "california", "colorado", "connecticut", "delaware", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", "kentucky","louisiana", "maine", "maryland", "massachusetts", "michigan", "minnesota", "mississippi", "missouri", "montana", "nebraska", "nevada", "new hampshire", "new jersey", "new mexico", "new york", "north carolina", "north dakota", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode island", "south carolina","south dakota", "tennessee", "texas", "utah", "vermont", "virginia", "washington", "west virginia", "wisconsin", "wyoming")

state_code <- data.frame(state_inits, state_names)
names(state_code) <- c("initials", "name")
```

```{r, label = "prepare datasets for merge"}
contribution_code <- contributions %>%
  select(cand_id, state) %>%
  group_by(state) %>%
  mutate(donor_state = state)

candidate_code <- candidates %>%
  mutate(party = cand_party_affiliation) %>%
  filter(party == "DEM" | party == "REP") %>%
  select(cand_id, party, cand_state)
```

```{r, label = "merge data and extract info", warning = FALSE}
contributions_state <- contribution_code %>%
  left_join(candidate_code, by = "cand_id") %>%
  filter(donor_state == cand_state) %>%
  select(-donor_state) %>%
  left_join(state_code, by = c("cand_state" = "initials")) %>%
  filter(!is.na(name)) %>%
  group_by(party, name) %>%
  summarize(sum = n(), num_cand = n_distinct(cand_id)) %>%
  mutate(don_cand = round(sum / num_cand)) %>%
  ungroup() %>%
  select(-sum, -num_cand)
```

```{r, label = "plot graphics", warning = FALSE, message = FALSE}
makeplot <- function(party1, color1) {
  
  states <- map_data("state")
  
  temp <- contributions_state %>% 
            filter(party == party1) %>%
            select(-party)
  
  temp2 <- states %>%
            left_join(temp, by = c("region" = "name")) %>%
            arrange(order)
  
  ggplot(data = temp2) +
    geom_polygon(aes(x = long, 
                     y = lat, 
                     fill = don_cand, 
                     group = region)) +
    coord_fixed(1.3) +
    scale_x_continuous(breaks = NULL) + 
    scale_y_continuous(breaks = NULL) +
    scale_fill_distiller(palette = color1, 
                         name = "Donations per Candidate", 
                         direction = 1, 
                         limits = c(0, 600),
                         breaks = c(0, 300, 600)) +
    theme(legend.position = "bottom", 
          panel.background = element_blank(),
          title = element_text(color= "#565151"), 
          axis.title = element_blank(),
          plot.title = element_text(face = "bold", size = 22, hjust = 0.5),
          text = element_text(family = "Baskerville"))
}
```

```{r, label = "graphics", warning = FALSE, message = FALSE}

makeplot("REP", "Reds") +
    labs(title = "2011 - 2012 Republican Donations", caption = "Source: Federal Election Commission")

makeplot("DEM", "Blues") +
    labs(title = "2011 - 2012 Democratic Donations", caption = "Source: Federal Election Commission")

```

<<<<<<< HEAD
=======
#General Trends
Donors contributed at higher rates in swing states, reflecting the understanding that money is power in elections. States that heavily lean towards one party had decreased donations in both parties because everyone in those states understand that donations will have less of an impact on the election compared to swing states. For that reason, donations will often flood into swing states from solid states because that is where money matters.

Filtering out donations from out of state reflects the same trend as when donations from out of state are included- the donation rate is higher in swing states regardless, pointing towards the fact that individuals in swing states are likely more politically involved because they have to fight for the political direction of the party more often than a resident of a solidly Democrat or Republican state.

#Romney Effect
While the general trend of donations corresponded with swing states, Massachusetts stands out in donations to Republican candidates. This is likely due to the fact that Mitt Romney, the Republican candidate for President, was the former governor of Massachusetts. Massachusetts is a strongly democratic state, but the fact that Romney was from Massachusetts would encourage a subset of the population to donate.

#Wyoming
Wyoming is an odd case, because there are no recorded democratic donations present in the FEC data. The absence of data can point to errors in recording or perhaps strong apathy in the Democratic party in the solidly Republican state.
>>>>>>> 6a76230f1a3e55f8dcaac629a18daf7f68f6c58a
```{r, label = "word count"}
devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)

#go to tools <- addin and click on word count
wordcountaddin:::text_stats()
```

